---
output:
  word_document: default
  html_document: default
---


```{r}
rm(list=ls())

setwd("C:\\Users\\thoma\\Documents\\UIC\\Courses\\IDS 572 Data Mining\\Asst4")

library(readxl)
library(dplyr)
library(tidyr)
library(randomForest)




Train_multiclass <- read_excel("IMB651-XLS-ENG.xlsx", sheet = "Training Data for Multi-Class M",
                    col_names = T)

Test_multiclass <- read_excel("IMB651-XLS-ENG.xlsx", sheet = "Test Data for Multi-Class Model",
                    col_names = T)

```


```{r}
print(dim(Train_multiclass))
print(dim(Test_multiclass))

```


```{r}

str(Train_multiclass)


# types of variables in the data
table(unlist(lapply(Train_multiclass, class)))

## Checking for missing values
sum(is.na(Train_multiclass))

## Target Variables
# NPS_Status

table(Train_multiclass$NPS_Status)


# dropping the 1st column
Train2_multiclass <- Train_multiclass[,-1]

```


```{r}
# Q 6 Part 1 
## Converting a 3 class problem to a Binary Class problem for Detractors

Train2_multiclass$Detractor_class <-ifelse(Train2_multiclass$NPS_Status=="Detractor",1,0)
table(Train2_multiclass$Detractor_class)

```



## Converting survey questions to ordinal variables
```{r}

cols_exclude <- c("CE_NPS","NPS_Status","AdmissionDate","DischargeDate",
                  "Estimatedcost","AgeYrs","HospitalNo2",
                  "LengthofStay","MaritalStatus","Sex",
                   "BedCategory","Department", "InsPayorcategory",
                   "State","Country","STATEZONE","Detractor_class")


## subsetting for only survey variables
survey_vars <- Train2_multiclass[, !(colnames(Train2_multiclass) %in% cols_exclude), drop                 = FALSE]

colnames(survey_vars)

## converting survey variables to ordinal variables
ordinal_vars <- survey_vars %>% 
  mutate_if(is.numeric,as.ordered)



cols_to_keep <- c("LengthofStay","MaritalStatus","Sex",
                  "BedCategory","Department", "InsPayorcategory",
                  "Estimatedcost","AgeYrs")

Train3_multiclass <- cbind(Train2_multiclass[,cols_to_keep],ordinal_vars,
                           Train2_multiclass["NPS_Status"])


```


## Prepping Test Data

```{r}
Test2_multiclass <- Test_multiclass[,-1]

## Creating a 2 class variable for Detractors

Test2_multiclass$Detractor_class <- ifelse(Test2_multiclass$NPS_Status=="Detractor",1,0)
table(Test2_multiclass$Detractor_class)

## subsetting for only survey variables
survey_vars_test <- Test2_multiclass[, !(colnames(Test2_multiclass) %in% cols_exclude),                       drop = FALSE]

colnames(survey_vars_test)

## converting survey variables to ordinal variables
ordinal_vars_test <- survey_vars_test %>% 
  mutate_if(is.numeric,as.ordered)


dim(ordinal_vars)


## Multiclass dataset
Test3_multiclass <- cbind(Test2_multiclass[,cols_to_keep],ordinal_vars_test,
                           Test2_multiclass["NPS_Status"])

```


##Removing variables for Quasi_complete Seperation

```{r}
colnames(Train3_multiclass)
quasi_vars <- c("MaritalStatus","BedCategory","LengthofStay")

Train4_multiclass <- Train3_multiclass[, !(colnames(Train3_multiclass) %in%                                   quasi_vars),drop = FALSE]

Test4_multiclass <- Test3_multiclass[, !(colnames(Test3_multiclass) %in%                                   quasi_vars),drop = FALSE]

```



## Q 7
```{r}

# Converting character columns to factor variables for Randomforest

class(Train4_multiclass$NPS_Status)

Train4_multiclass$NPS_Status <- as.factor(Train4_multiclass$NPS_Status)

Train5_multiclass <- Train4_multiclass %>% 
  mutate_if(is.character,as.factor)


# Converting character columns to factor variables for Randomforest

Test4_multiclass$NPS_Status <- as.factor(Test4_multiclass$NPS_Status)

Test5_multiclass <- Test4_multiclass %>% 
  mutate_if(is.character,as.factor)


## Adaboosting



```{r}

library(adabag)

Test6_multiclass <- Test5_multiclass %>% 
                    filter(NS_NURSEPROACTIVENESS!=1)
  
dim(Test5_multiclass)
dim(Test6_multiclass)
library(rpart)

maxdep <- c(1:5)
pr_val_err <- matrix()


for(i in maxdep){
    
    boost_model <- boosting(NPS_Status~., data = Train5_multiclass, boos = T,
                        mfinal = 100, coeflearn = "Breiman",
                        control=rpart.control(maxdepth=i))
    
    boost_pred <- predict.boosting(boost_model,newdata=Test6_multiclass)
    
    # pr_val_err[i] <- which.min(boost_pred$error)
    pr_val_err[i] <- boost_pred$error
    print(paste0(maxdep[i],",",pr_val_err[i]))
    

  }

minsplit <- c(5,10,15)
pr_val_err2 <- matrix()

for(j in 1:length(minsplit)){
    
    boost_model <- boosting(NPS_Status~., data = Train5_multiclass, boos = T,
                        mfinal = 100, coeflearn = "Breiman",
                        control=rpart.control(maxdepth=4, minsplit = j))
    
    boost_pred <- predict.boosting(boost_model,newdata=Test6_multiclass)
    
    # pr_val_err[i] <- which.min(boost_pred$error)
    pr_val_err2[j] <- boost_pred$error
    print(paste0(minsplit[j],",",pr_val_err2[j]))
    

}


trees<- c(100,200,300,400)
pr_val_err3 <- matrix()

for(k in 1:length(trees)){
    
    boost_model <- boosting(NPS_Status~., data = Train5_multiclass, boos = T,
                        mfinal = k, coeflearn = "Breiman",
                        control=rpart.control(maxdepth=4, minsplit = 10,cp = 0.01))
    
    boost_pred <- predict.boosting(boost_model,newdata=Test6_multiclass)
    
    # pr_val_err[i] <- which.min(boost_pred$error)
    pr_val_err3[k] <- boost_pred$error
    print(paste0(trees[k],",",pr_val_err3[k]))
    

  }




set.seed(123)

boost_cv<- boosting.cv(NPS_Status~., data = Train5_multiclass, v = 5, boos = TRUE,mfinal = 300, coeflearn = "Breiman",  control=rpart.control(maxdepth=4, cp =0.01, minsplit =10), par=FALSE)

boost_cv$error
boost_cv$confusion

set.seed(123)
boost2 <- boosting(NPS_Status~., data = Train5_multiclass, boos = T,
                        mfinal = 300, coeflearn = "Breiman",
                   control=rpart.control(maxdepth=4, cp =0.01, minsplit =10))


adaboost_pred <- predict.boosting(boost2,newdata=Test6_multiclass)
# adaboost_pred

cf <- adaboost_pred$confusion
cf
boost_err <- adaboost_pred$error

acc <-  1-boost_err
acc

sens_Class1 <- cf[1,1]/(cf[1,1]+cf[1,2]+cf[1,3])
sens_Class1

sens_Class2 <- cf[2,2]/(cf[2,1]+cf[2,2]+cf[2,3])
sens_Class2

sens_Class3 <- cf[3,3]/(cf[3,1]+cf[3,2]+cf[3,3])
sens_Class3



```



```{r}

## Undersampling
library(caret)
under <- downSample(Train5_multiclass[,-41], Train5_multiclass$NPS_Status,
           list = F, yname = "NPS_Status")
class(under)
dim(Train5_multiclass)
dim(under)
table(Train5_multiclass$NPS_Status)

table(under$NPS_Status)

boost_under <- boosting(NPS_Status~., data = under, boos = T,
                        mfinal = 300, coeflearn = "Breiman",
                   control=rpart.control(maxdepth=4, cp =0.01, minsplit =10))


adaboost_pred_under <- predict.boosting(boost_under,newdata=Test6_multiclass)
# adaboost_pred

cf2 <- adaboost_pred_under$confusion
cf2
boost_err2 <- adaboost_pred_under$error

acc2 <-  1-boost_err2
acc2

sens_Class1 <- cf2[1,1]/(cf2[1,1]+cf2[1,2]+cf2[1,3])
sens_Class1

sens_Class2 <- cf2[2,2]/(cf2[2,1]+cf2[2,2]+cf2[2,3])
sens_Class2

sens_Class3 <- cf2[3,3]/(cf2[3,1]+cf2[3,2]+cf2[3,3])
sens_Class3


```


## Oversampling

```{r}
over <- upSample(Train5_multiclass[,-41], Train5_multiclass$NPS_Status,
           list = F, yname = "NPS_Status")
class(under)
dim(Train5_multiclass)
dim(over)
table(Train5_multiclass$NPS_Status)

table(over$NPS_Status)

boost_over <- boosting(NPS_Status~., data = over, boos = T,
                        mfinal = 300, coeflearn = "Breiman",
                   control=rpart.control(maxdepth=4, cp =0.01, minsplit =10))


adaboost_pred_over <- predict.boosting(boost_over,newdata=Test6_multiclass)
# adaboost_pred

cf3 <- adaboost_pred_over$confusion
cf3
boost_err3 <- adaboost_pred_over$error

acc3 <-  1-boost_err3
acc3

sens_Class1 <- cf3[1,1]/(cf3[1,1]+cf3[1,2]+cf3[1,3])
sens_Class1

sens_Class2 <- cf3[2,2]/(cf3[2,1]+cf3[2,2]+cf3[2,3])
sens_Class2

sens_Class3 <- cf3[3,3]/(cf3[3,1]+cf3[3,2]+cf3[3,3])
sens_Class3

```

## SMOTE

```{r}
library(DMwR)
# Train5_multiclass$NPS_Status <- as.character.factor(Train5_multiclass$NPS_Status)
Train5_multiclass$NPS_Status <- factor(Train5_multiclass$NPS_Status)


Train5_multiclass <- as.data.frame(Train5_multiclass)
sum(is.na(Train5_multiclass))
sum(is.na(data_smote))
dim(Train5_multiclass)
dim(data_smote)

data_smote <- DMwR::SMOTE(NPS_Status~., data = Train5_multiclass
                    , perc.over = 500, k =10, 
                    perc.under = 100)

data_smote2 <- data_smote[complete.cases(data_smote), ]
dim(data_smote2)



table(under$NPS_Status)

boost_smote <- boosting(NPS_Status~., data = data_smote2, boos = T,
                        mfinal = 300, coeflearn = "Breiman",
                   control=rpart.control(maxdepth=4, cp =0.01, minsplit =10))


adaboost_pred_smote <- predict.boosting(boost_smote,newdata=Test6_multiclass)
# adaboost_pred

cf4 <- adaboost_pred_smote$confusion
cf4
boost_err2 <- adaboost_pred_smote$error

acc4 <-  1-boost_err2
acc4

sens_Class1 <- cf4[1,1]/(cf4[1,1]+cf4[1,2]+cf4[1,3])
sens_Class1

sens_Class2 <- cf4[2,2]/(cf4[2,1]+cf4[2,2]+cf4[2,3])
sens_Class2

sens_Class3 <- cf4[3,3]/(cf4[3,1]+cf4[3,2]+cf4[3,3])
sens_Class3


```

